\documentclass[12pt,t]{beamer}
\usepackage{graphicx}
\setbeameroption{hide notes}
\setbeamertemplate{note page}[plain]
\usepackage{listings}

\input{header.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% end of header
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% title info
\title{Reproducible Research}
\author{\href{http://kbroman.org}{Karl Broman}}
\institute{Biostatistics \& Medical Informatics, UW{\textendash}Madison}
\date{\href{http://kbroman.org}{\tt \scriptsize \color{foreground} kbroman.org}
\\[-4pt]
\href{https://github.com/kbroman}{\tt \scriptsize \color{foreground} github.com/kbroman}
\\[-4pt]
\href{https://twitter.com/kwbroman}{\tt \scriptsize \color{foreground} @kwbroman}
}


\begin{document}

% title slide
{
\setbeamertemplate{footline}{} % no page number here
\frame{
  \titlepage
  \note{Blah blah blah reproducible blah blah.}
} }


\begin{frame}[c]{}


\centering
\Large

Reproducible

\bigskip

\onslide<2->{{\color{lolit} vs.}}

\bigskip

\only<1|handout 0>{{\color{background} invisible text}}
\only<2>{Replicable}
\only<3 | handout 0>{Correct}

\note{Computational work is
  {\color{nhilit} reproducible} if one can take the data and code and produce
  the same set of results. {\color{nhilit} Replicable} is more stringent: can
  someone repeat the experiment and get the same results?

  Reproducibility is a minimal standard. That something is
  reproducible doesn't imply that it is correct. The code may have bugs. The
  methods may be poorly behaved. There could be experimental
  artifacts.

  (But reproducibility is probably correlated with correctness.)

  Note that some scientists say replicable for what I call
  reproducible, and vice versa.
}
\end{frame}



\begin{frame}{Levels of quality}


\vspace{24pt}

\bi
\itemsep12pt
\item Are the tables and figures reproducible from the code and data?
\item Does the code actually do what you think it does?
\item In addition to {\color{hilit} what} was done, is it clear
  {\color{hilit} why} it was done?
  \bi
  \item[] (e.g., how were parameter settings chosen?)
  \ei
\item Can the code be used for other data?
\item Can you extend the code to do other things?
\ei

\note{Reproducibility is not black and white. And the ideal is hard to
  achieve.
}
\end{frame}



\begin{frame}{Basic principles}


\vspace{24pt}

\bi
\itemsep12pt
\item Everything via code
\item Everything automated
    \bi
    \item[] Workflow and dependencies clearly documented
    \ei
\item Get the data in the most-raw form possible
\item Get any/all data and meta-data possible
\item Keep track of the {\color{hilit} provenance} of all data files
\item Be self-sufficient
\ei

\note{Pointing and clicking is not reproducible. Ideally, you press
  just one button.

Make sure you have all of the data and that you know exactly where it
came from.

But what is {\color{nhilit} raw} data? How far back should you go?
Data that I get from collaborators has usually gone through a
considerable amount of pre-processing. Should we have captured that,
in order for the work to be considered reproducible?

If your collaborator asks, ``In what form would you like the data?''
you should respond, ``In its current form.''
}
\end{frame}

\begin{frame}{Why do we care?}

\vspace{24pt}

\bi
\itemsep12pt
\item Avoid embarrassment
\item More likely correct
\item Save time, in the long run
\item Greater potential for extensions; higher impact
\ei

\note{Doing things properly (writing clear, documented, well-tested
  code) is time consuming, but it could save you a ton of aggravation
  down the road.  Ultimately, you'll be more efficient, and your work
  will have greater impact.

Your code and analyses will be easier to debug, maintain, and extend.
}
\end{frame}


\begin{frame}[c]{}

\centering
\large
Your closest collaborator is you six months ago, but you
don't reply to emails.
\note{I heard this from Paul Wilson, UW-Madison.
}
\end{frame}


\begin{frame}{What could go wrong?}

\vspace{24pt}

\bi
\itemsep12pt
\item "The attached is similar to the code we used."
\item "Where did this data file come from?!"
\item "Can you repeat the analysis, omitting subject X?"
\item "This part of your script is now giving an error."
\ei

\note{If you've not heard any of these things, it's just a matter of
  time.
}
\end{frame}


\begin{frame}{Need to avoid}


\vspace{24pt}

\bi
\itemsep12pt
\item Open a file to extract as CSV
\item Open a data file to do even a slight edit
\item Paste results into the text of a manuscript
\item Copy-paste-edit tables
\item Copy-paste-adjust figures
\ei

\note{If you do anything ``by hand'' once, you'll have to do it 100
  times.
}
\end{frame}

\begin{frame}{Basic tools}


\vspace{24pt}

\bi
\itemsep12pt
\item Automation with Make
\item Unix command line
\item Latex and Markdown
\item Knitr
\item Version control with git
\item R packages
\item Python (or Ruby or Perl)
\ei

\note{These are the basic tools that I think are important for
  reproducible computational research; they form the core topics for
  the course.

  Make is for automation and for documenting dependencies. For
  reproducibility, the command line is your best friend. Latex and
  Markdown allow preparation of beautiful documents without pointing
  or clicking. Knitr is for combining code and text; knitr and make
  are the key tools for reproducibility. Version control isn't
  strictly necessary for reproducibility, but once you get the hang of
  it, you'll never go back. R's packaging system is among its best
  features. A scripting language like Python is invaluable for
  manipulating data files. Many things that are awkward in R are easy
  in Python.
}
\end{frame}


\begin{frame}{Other topics}

\vspace{24pt}

\bi
\itemsep12pt
\item Organizing projects
\item Writing clear code
\item Don't Repeat Yourself (DRY)
\item Testing and debugging
\item Handling big jobs
\item Licenses; human subjects data
\ei

\note{We'll also cover all of these things.

The organization of the data and code for a project is a major
determinant of whether others will be able to make sense of it.
Good code is not just correct but is clearly written. Code is easier
to maintain and understand if it is modular. Adding good tests will
help you to find problems in your code earlier rather than later. But
you'll spend a lot of time debugging, so we should talk about
debugging strategies. Big computational jobs (particularly big
computer simulations) raise additional issues; reproducibility is
especially tricky. Finally, code that you distribute should be
licensed. And if you're working with data on human subjects, you
need to be extra careful.
}
\end{frame}


\begin{frame}{Don't Repeat Yourself}

\vspace{24pt}

\bi
\itemsep12pt
\item In code, in documentation, etc.
\item Repeated bits of code are harder to maintain
    \bi
    \item[] Write a function
    \ei
\item Use documentation systems like Roxygen2
    \bi
    \item[] Documentation in just one place
    \ei
\item Make use of others' code
\ei

\note{DRY is among the more important concepts/techniques.

For example, I helped organize a meeting
in Madison in 2013. The program book and website both drew
information from a single basic source. Change that one document and
both the program and web site are updated.

My R/qtl package is an {\color{nhilit} anti}-example.
}
\end{frame}


\end{document}
